---
title: "R Notebook"
---

```{r}
library(dplyr)
library(terra)
library(biomod2)
library(pROC)
source("bm_ROC.R")

if(!dir.exists("inputs")){dir.create("inputs")}
if(!dir.exists("outputs")){dir.create("outputs")}
grid <- rast("../../../grid/outputs/grid_250m.tif")
start_time <- Sys.time()
```

## Settings
## Species: "Abra alba", "Bathyporeia pilosa", "Macoma balthica", "Marenzelleria neglecta"
```{r}
spec <- "Bathyporeia pilosa"
spec_ <- gsub(" ", "_", spec)
mod_methods <- c("XGBOOST", "RF", "GBM")
sdir <- paste0("outputs/", gsub(" ", "_", spec))
if(!dir.exists(sdir)){dir.create(sdir)}

mod_id <- as.character(sample(1:999999999, 1))
resp_n <- paste0("temp", mod_id)
proj_n <- "misc"
```

## Import observations
```{r}
#mod_dir <- paste0(sdir, "/", gsub(" ", "_", spec), "_model.sdm")
obs_dir <- paste0("../input_data/observations/", spec_, "_observations.csv")
obs <- read.csv(obs_dir)
train <- select(obs, train1, train2, train3)
train_base <- train
obs <- select(obs, count, long, lat)
obs <- vect(obs, geom = c("long", "lat"), crs="+proj=longlat +datum=WGS84")
obs <- project(obs, crs(grid))
```

## Import predictors
```{r}
l <- list.files("../input_data/predictors", full.names = T, pattern = "\\.tif$")
preds <- rast(l)
names(preds) <- tools::file_path_sans_ext(basename(l))

# Aggregate predictors for faster testing
preds <- aggregate(preds, fact = 5, fun = "max", na.rm = T)
```

## Prepare model
```{r}
dat <- BIOMOD_FormatingData(resp.var = obs,
                            expl.var = preds,
                            resp.name = resp_n,
                            dir.name = "outputs")

dat
plot(dat)
```

## Specify training/testing
```{r}
colnames(train) <- c("_allData_RUN1", "_allData_RUN2", "_allData_RUN3")
train <- train %>% mutate_all(~as.logical(.))
mod_cv <- bm_CrossValidation(bm.format = dat,
                             strategy = 'user.defined',
                             user.table = train)
```
## Single models
```{r}
mod <- BIOMOD_Modeling(bm.format = dat,
                       models = mod_methods,
                       modeling.id = mod_id,
                       CV.strategy = 'user.defined',
                       CV.user.table = mod_cv,
                       OPT.strategy = 'bigboss',
                       var.import = 2,
                       metric.eval = c('TSS','ROC'),
                       do.progress = FALSE)
```

## Tuning (if desired)
```{r}
#OptionsBigboss@models
#
#opt.d <- bm_ModelingOptions(data.type = 'binary',
#                            models = mod_methods,
#                            strategy = 'bigboss')
#
#RF_tuned <- bm_Tuning(model = 'RF',
#             tuning.fun = 'rf', ## see in ModelsTable
#             do.formula = FALSE,
#             bm.options = opt.d@options$RF.binary.randomForest.randomForest,
#             bm.format = dat,
#             calib.lines = get_calib_lines(mod))
#
#ANN_tuned <- bm_Tuning(model = 'ANN',
#             tuning.fun = 'avNNet', ## see in ModelsTable
#             do.formula = FALSE,
#             bm.options = opt.d@options$ANN.binary.nnet.nnet,
#             bm.format = dat,
#             calib.lines = get_calib_lines(mod))
```

## Limited tuning
```{r}
#opt.d <- bm_ModelingOptions(data.type = 'binary',
#                            models = mod_methods,
#                            strategy = 'bigboss')
#
#XG_tuned <- bm_Tuning(model = 'XGBOOST',
#                      tuning.fun = 'xgbTree', ## see in ModelsTable
#                      do.formula = FALSE,
#                      bm.options = opt.d@options$XGBOOST.binary.xgboost.xgboost,
#                      params.train = list(XGBOOST.nrounds = c(10, 20, 50, 100),
#                                          XGBOOST.max_depth = c(1, 6)),
#                      bm.format = dat,
#                      calib.lines = get_calib_lines(mod))
#
#ANN_tuned <- bm_Tuning(model = 'ANN',
#                      tuning.fun = 'avNNet', ## see in ModelsTable
#                      do.formula = FALSE,
#                      bm.options = opt.d@options$ANN.binary.nnet.nnet,
#                      params.train = list(ANN.size = c(2, 6, 8)),
#                      bm.format = dat,
#                      calib.lines = get_calib_lines(mod))
```


## Alter model parameters (if desired)
```{r}
opt.b <- bm_ModelingOptions(data.type = 'binary',
                            models = mod_methods,
                            strategy = 'bigboss')

#user.RF <- list('_allData_allRun' = list(ntree = 250))
#user.ANN <- list('_allData_allRun' = list(size = 16))
#user.XGBOOST <- list('_allData_allRun' = list(nrounds = 10))

user.RF <- vector("list", 3)
names(user.RF) <- paste0("_allData_RUN", 1:3)
user.RF <- lapply(user.RF, function(x) list(ntree = 250))

user.ANN <- vector("list", 3)
names(user.ANN) <- paste0("_allData_RUN", 1:3)
user.ANN <- lapply(user.ANN, function(x) list(size = 6, decay = 0.2, maxit = 1000))

user.XGBOOST <- vector("list", 3)
names(user.XGBOOST) <- paste0("_allData_RUN", 1:3)
user.XGBOOST <- lapply(user.XGBOOST, function(x) list(nrounds = 50))

user_val <- list(RF.binary.randomForest.randomForest = user.RF,
                 ANN.binary.nnet.nnet = user.ANN,
                 XGBOOST.binary.xgboost.xgboost = user.XGBOOST)
```


```{r}
user_opt <- bm_ModelingOptions(data.type = 'binary',
                               models = mod_methods,
                               strategy = "user.defined",
                               user.val = user_val,
                               user.base = "bigboss",
                               bm.format = dat, 
                               calib.lines = get_calib_lines(mod))

mod_tuned <- BIOMOD_Modeling(bm.format = dat,
                             modeling.id = paste0(mod_id, "_tuned"),
                             models = mod_methods,
                             CV.strategy = 'user.defined',
                             CV.user.table = mod_cv,
                             OPT.user = user_opt,
                             metric.eval = c('TSS','ROC'),
                             var.import = 2)
```

## Compare tuned and un-tuned models
```{r}
bm_PlotEvalMean(mod, do.plot = FALSE)
bm_PlotEvalMean(mod_tuned, do.plot = FALSE)
```

## Plot ROC curves
```{r}
mod_select <- mod # Choose tuned or untuned model
train <- get_calib_lines(mod_select)
if(any(colnames(train) == "_allData_allRun")){
  train <- train[ , -which(colnames(train) %in% c("_allData_allRun"))]
}
cv_runs <- ncol(train)

# Export ROC curves
pdf(file = paste0(sdir, "/", gsub(" ", "_", spec), "_ROC_curves.pdf"),
    width = cv_runs*3,
    height = length(mod_methods)*3)

bm_ROC(mod_select, mod_methods)
dev.off()

mod_select <- mod_tuned # Choose tuned or untuned model
train <- get_calib_lines(mod_select)
if(any(colnames(train) == "_allData_allRun")){
  train <- train[ , -which(colnames(train) %in% c("_allData_allRun"))]
}
cv_runs <- ncol(train)

# Export ROC curves
pdf(file = paste0(sdir, "/", gsub(" ", "_", spec), "_ROC_curves_tuned.pdf"),
    width = cv_runs*3,
    height = length(mod_methods)*3)

bm_ROC(mod_select, mod_methods)
dev.off()
```

## Select which model to use
```{r}
mod <- mod_tuned
```

## Performance
```{r}
# Export performance metrics
mod_eval <- get_evaluations(mod)
mod_eval <- mod_eval %>%
  select(algo, run, metric.eval, calibration, validation)
colnames(mod_eval) <- c("algo", "cv_run", "metric", "calibration", "validation")
write.csv(mod_eval, paste0(sdir, "/", spec_, "_individial_model_performance.csv"), row.names = FALSE)

png(file = paste0(sdir, "/", spec_, "_evaluation_plot.png"))
bm_PlotEvalMean(mod)
dev.off()
```

## Ensemble model
```{r}
emod <- BIOMOD_EnsembleModeling(bm.mod = mod,
                                models.chosen = 'all',
                                em.by = 'PA+run',
                                em.algo = c('EMwmean'),
                                metric.select = c('ROC'),
                                metric.select.thresh = c(0.7),
                                metric.eval = c('TSS', 'ROC'),
                                var.import = 2,
                                EMci.alpha = 0.05,
                                EMwmean.decay = 'proportional')
```

## Ensemble performance
```{r}
emod_eval <- get_evaluations(emod)
emod_eval <- emod_eval %>%
  select(algo, merged.by.run, metric.eval, calibration, validation)
emod_eval$algo <- "ensemble"
colnames(emod_eval) <- c("algo", "cv_run", "metric", "calibration", "validation")
write.csv(emod_eval, paste0(sdir, "/", spec_, "_ensemble_model_performance.csv"), row.names = FALSE)

# Tjur's R2
true_obs <- as.logical(obs$count)

tjur_df <- data.frame("algo" = "ensemble", "run" = paste0("RUN", 1:3), "tjurR2" = NA)

for(i in 1:3){
  r <- paste0("RUN", i)
  emod_preds <- get_predictions(emod) %>%
    filter(merged.by.run == r)
  mt <- mean(emod_preds$pred[which(true_obs)])/1000
  mf <- mean(emod_preds$pred[which(!true_obs)])/1000
  
  tjur_df$tjurR2[i] <- mt - mf 
}

write.csv(tjur_df, paste0(sdir, "/", spec_, "_ensemble_tjurR2.csv"), row.names = FALSE)
```


## Ensemble model ROC curves
```{r}
pdf(file = paste0(sdir, "/", gsub(" ", "_", spec), "_ensemble_ROC_curves.pdf"),
    width = cv_runs*3,
    height = 3)

bm_ROC(mod, mod_methods, emod)
dev.off()
```


## Response curves
```{r}
bm_plot <- bm_PlotResponseCurves(bm.out = emod, 
                      models.chosen = get_built_models(emod),
                      fixed.var = 'median',
                      do.progress = FALSE,
                      do.plot = FALSE)

pdf(file = paste0(sdir, "/", spec_, "_response_curves.pdf"),
    width = 20,
    height = 20)

bm_plot[["plot"]]

dev.off()
```

## Predictions
```{r}
mod_proj <- BIOMOD_Projection(bm.mod = mod,
                                  proj.name = proj_n,
                                  new.env = preds,
                                  models.chosen = 'all',
                                  metric.binary = 'all',
                                  metric.filter = 'all',
                                  build.clamping.mask = T)

plot(mod_proj)
```

```{r}
emod_proj <- BIOMOD_EnsembleForecasting(bm.em = emod,
                                        bm.proj = mod_proj,
                                        models.chosen = 'all',
                                        metric.binary = 'all',
                                        metric.filter = 'all'
                                        )
plot(emod_proj)
```

## Export relevant files
```{r}
# Model files
dir <- paste0("outputs/", spec_, "/models")
if(!dir.exists(dir)){dir.create(dir)}
mod_dir <- paste0("outputs/", resp_n, "/models/", mod_id)
file.copy(list.files(mod_dir, full.names = T), dir, overwrite = TRUE)

# Individual model projections
mod_stack <- terra::unwrap(mod_proj@proj.out@val)
n <- names(mod_stack)
p <- paste0(resp_n, "_allData")
n <- gsub(p, spec_, n)
names(mod_stack) <- n
dir <- paste0("outputs/", spec_, "/individual_projections")
if(!dir.exists(dir)){dir.create(dir)}
for(i in 1:nlyr(mod_stack)){
  writeRaster(mod_stack[[i]], filename = paste0(dir, "/", names(mod_stack)[i], ".tif"), overwrite = T)
}

# Ensemble model projections
mod_stack <- terra::unwrap(emod_proj@proj.out@val)
n <- names(mod_stack)
n <- gsub(resp_n, spec_, n)
names(mod_stack) <- n
dir <- paste0("outputs/", spec_, "/ensemble_projections")
if(!dir.exists(dir)){dir.create(dir)}
for(i in 1:nlyr(mod_stack)){
  writeRaster(mod_stack[[i]], filename = paste0(dir, "/", names(mod_stack)[i], ".tif"), overwrite = T)
}

# Observations, CV set & predictions
predictions <- get_predictions(emod, model.as.col = T, algo = 'EMwmean')
obs <- read.csv(paste0("../input_data/observations/", spec_, "_observations.csv"))
train <- select(obs, train1, train2, train3)
pa_predictions <- predictions
cut <- get_evaluations(emod, metric.eval = "TSS")

for(i in 1:3){
  pa_predictions[,i] <- ifelse(pa_predictions[,i] > cut$cutoff[i], 1, 0)
}

df <- data.frame("true_observation" = obs$count,
                 "train1" = train$train1,
                 "train2" = train$train2,
                 "train3" = train$train3,
                 "predictions_train1" = predictions[,1]/1000,
                 "predictions_train2" = predictions[,2]/1000,
                 "predictions_train3" = predictions[,3]/1000,
                 "pa_predictions_train1" = pa_predictions[,1],
                 "pa_predictions_train2" = pa_predictions[,2],
                 "pa_predictions_train3" = pa_predictions[,3]
                 )

write.csv(obs, paste0("outputs/", spec_, "/", spec_, "_observations.csv"), row.names = F)
write.csv(train, paste0("outputs/", spec_, "/", spec_, "_CVset.csv"), row.names = F)
write.csv(df, paste0("outputs/", spec_, "/", spec_, "_predictions.csv"), row.names = F)

# Record run time
stop_time <- Sys.time()
run_time <- difftime(stop_time, start_time, units = "hours")
run_df <- data.frame(species = spec, run_time = run_time, models = paste(mod_methods, collapse = ","))
write.csv(run_df, paste0(sdir, "/", spec_, "_run_time.csv"), row.names = FALSE)

# Delete biomod2 created directory
dir <- paste0("outputs/", resp_n)
unlink(dir, recursive = T)
```


```{r}
# Model files
dir <- paste0("outputs/", spec_, "/models")
if(!dir.exists(dir)){dir.create(dir)}
mod_dir <- paste0("outputs/", resp_n, "/models/", mod_id)
file.copy(list.files(mod_dir, full.names = T), dir, overwrite = TRUE)

# Observations, CV set & predictions
#obs <- get_formal_data(mod)@data.species
#train <- get_calib_lines(mod)
#predictions <- get_predictions(mod, model.as.col = T, algo = mod_methods)

predictions <- get_predictions(emod, model.as.col = T, algo = 'EMwmean')
obs <- read.csv(paste0("../input_data/observations/", spec_, "_observations.csv"))
train <- select(obs, train1, train2, train3)
pa_predictions <- predictions
cut <- get_evaluations(emod, metric.eval = "TSS")

for(i in 1:3){
pa_predictions[,i] <- ifelse(pa_predictions[,i] > cut$cutoff[i], 1, 0)
}

df <- data.frame("true_observation" = obs$count,
                 "train1" = train$train1,
                 "train2" = train$train2,
                 "train3" = train$train3,
                 "predictions_train1" = predictions[,1]/1000,
                 "predictions_train2" = predictions[,2]/1000,
                 "predictions_train3" = predictions[,3]/1000,
                 "pa_predictions_train1" = pa_predictions[,1],
                 "pa_predictions_train2" = pa_predictions[,2],
                 "pa_predictions_train3" = pa_predictions[,3]
                 )

write.csv(obs, paste0("outputs/", spec_, "/", spec_, "_observations.csv"), row.names = F)
write.csv(train, paste0("outputs/", spec_, "/", spec_, "_CVset.csv"), row.names = F)
write.csv(df, paste0("outputs/", spec_, "/", spec_, "_predictions.csv"), row.names = F)

# Projections/full predictions
dir <- paste0("outputs/", spec_, "/", spec_, "_single_proj.tif")
mod_dir <- paste0("outputs/", resp_n, "/proj_", proj_n, "/proj_", proj_n, "_", resp_n, ".tif")
file.copy(mod_dir, dir, overwrite = TRUE)

dir <- paste0("outputs/", spec_, "/", spec_, "_ensemble_proj.tif")
mod_dir <- paste0("outputs/", resp_n, "/proj_", proj_n, "/proj_", proj_n, "_", resp_n, "_ensemble.tif")
file.copy(mod_dir, dir, overwrite = TRUE)

# Record run time
stop_time <- Sys.time()
run_time <- difftime(stop_time, start_time, units = "hours")
run_df <- data.frame(species = spec, run_time = run_time, models = paste(mod_methods, collapse = ","))
write.csv(run_df, paste0(sdir, "/", spec_, "_run_time.csv"), row.names = FALSE)

# Delete biomod2 created directory
dir <- paste0("outputs/", resp_n)
unlink(dir, recursive = T)
```


```{r}
bm_ROC(mod, mod_methods, emod)
```


